---
title: "hands-on-exercise-6"
description: |
  Spatial Weights - Neighbour Analysis
author:
  - name: Xu Pengtai
    url: https://www.linkedin.com/in/xupengtai/
date: 09-20-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Getting Started
- create a list of R packages required

```{r echo=TRUE, eval=TRUE}
packages <- c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages) {
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
```

## Import shapefile into r environment
```{r echo=TRUE, eval=TRUE}
hunan <- st_read(dsn = "data/shapefile", 
                 layer = "Hunan")
```

## Import csv file into r environment
```{r echo=TRUE, eval=TRUE}
hunan2012 <- read_csv("data/attribute/Hunan_2012.csv")
```

## Performing relational join
```{r echo=TRUE, eval=TRUE}
hunan <- left_join(hunan,hunan2012)
```

## Visualising regional development indicator
```{r echo=TRUE, eval=TRUE}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size = 0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol = 2)

```

## Computing (QUEEN) contiguity based neighbours
```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

- Get neighbour info of a particular county
```{r}

# get neighbour
wm_q[[1]]

# get county name
hunan$County[1]

# get neighbour county name
hunan$County[c(2,3,4,57,85)]

```

## Computing (ROOK) contiguity based neighbours
```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

## Visualising contiguity weights

- create coordinates

* note: longitude before latitude (x before y)
```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
coords <- cbind(longitude, latitude)
head(coords)
```

- Plotting Queen contiguity based neighbours maps
```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```
- Comparing Queen and Rook methods
```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")
```

## Computing distance based neighbours

### Determine the cut-off distance
- make use of knearneigh() of spdep

```{r}
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

- Computing fixed distance weight matrix
```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

- Another way to display the structure of the weight matrix is to combine table() and card() of spdep.
```{r}
table(hunan$County, card(wm_d62))
```

- The red lines show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.
```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")
```

## Computing adaptive distance weight matrix

- adaptive distance means fixed number of neighbours in this case
```{r}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
```

```{r}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## Weights based on IDW

- based on the inverse distance method (longer distance -> lower weights)
```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## Row-standardised weights matrix
- 'W': each neighbouring polygon assigned the same weight (may be biased against polygons on the edge)
- 'zero policy': allows for lists of non=neighbours

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```
## Application of Spatial Weight Matrix

### Spatial lag with row-standardized weights
- average of neighbouring polygons
```{r}

# calculate lag
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")

# join back to df
hunan <- left_join(hunan,lag.res)

# plot
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)
```



### Spatial lag as a sum of neighbouring values
```{r}
knn6a <- knn6
include.self(knn6a)
binary.knn6 <- lapply(knn6a, function(x) 0*x+1)
binary.knn6[1]
wm_knn6 <- nb2listw(knn6a, glist = binary.knn6, style = "B")
lag_knn6 <- lag.listw(wm_knn6, hunan$GDPPC)
lag.list.knn6 <- list(hunan$NAME_3, lag.listw(wm_knn6, hunan$GDPPC))
lag_knn6.res <- as.data.frame(lag.list.knn6)
colnames(lag_knn6.res) <- c("NAME_3", "lag_sum GDPPC")
hunan <- left_join(hunan, lag_knn6.res)
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```



Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


