---
title: "In-class_Ex05r1"
description: |
  Spatial Point Pattern Analysis
author:
  - name: Xu Pengtai
    url: https://www.linkedin.com/in/xupengtai/
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---

``` {r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      fig.retina = 3)
```

## Getting Started
- create a list of R packages required

```{r echo=TRUE, eval=TRUE}
packages <- c('maptools', 'sf', 'raster', 'spatstat', 'tidyverse', 'tmap', 'plotly', 'ggthemes')
for (p in packages) {
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
```


## Importing the Geospatial Data
```{r}
mpsz_sf <- st_read(dsn = 'data/shapefile',
                   layer = "MP14_SUBZONE_WEB_PL")
```

## Importing aspatial Data
```{r}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```

## Converting aspatial to geospatial
```{r}
CHAS_sf <- st_as_sf(CHAS, 
                    coords = c("X_COORDINATE",
                               "Y_COORDINATE"),
                    crs = 3414)  #1: 3414?
```

Note: *st_as_sf()* accepts coordinates in character data type.

```{r}
childcare$Lat <- as.numeric(childcare$Lat)
childcare$Lng <- as.numeric(childcare$Lng)
childcare_sf <- st_as_sf(childcare, 
                    coords = c("Lng",
                               "Lat"),
                    crs = 4326) %>%
  st_transform(crs=3414) 
```


## Plottinng for reviewinng

```{r}
tmap_mode("view")
tm_shape(childcare_sf) + 
  tm_dots(alpha = 0.4,
          col = "blue",
          size = 0.05)

```


## Geospatial Data Wrangling

### Converting from sf to Spatial* data frame

```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

### Converting Spatial data frame into spatial objects

as.SpatialPoint() or as.SpatialPolygon() of maptools package

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

### Converting from Spatial* objects into ppp objects

tmap does not understand ppp format

``` {r}
childcare_ppp <- as(childcare, "ppp")
CHAS_ppp <- as(CHAS, "ppp")
```

### Removing duplicate points using jitter

``` {r}
childcare_ppp_jitt <- rjitter(childcare_ppp,
                              retry = TRUE,
                              nsim = 1,
                              drop = TRUE)
any(duplicated(childcare_ppp_jitt))
```

``` {r}
CHAS_ppp_jitt <- rjitter(CHAS_ppp,
                              retry = TRUE,
                              nsim = 1,
                              drop = TRUE)
any(duplicated(CHAS_ppp_jitt))
```

### Extracting Punggol Planning Area

``` {r}
pg <- mpsz[mpsz@data$PLN_AREA_N=="PUNGGOL",]
```

### Converting SpatialPolygonsDataFrame into SpatialPolygonsObject

``` {r}
pg_sp <- as(pg, "SpatialPolygons")
```

### Converting SpatialPolygonsObject into owin object

``` {r}
pg_owin <- as(pg_sp, "owin")
```

### Extracting Spatial Points within the owin

``` {r}
childcare_pg <- childcare_ppp_jitt[pg_owin]
CHAS_pg <- CHAS_ppp_jitt[pg_owin]
```

```{r}
plot(childcare_pg)
```

### Pairwise Distance: L function

max distance until sign of clustering shown

```{r}

L_childcare <- envelope(childcare_pg, 
                        Lest,
                        nsim=99,
                        rank=1,
                        global = TRUE)

title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_childcare)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


