---
title: "hands-on-exercise-9"
description: |
  Geographically Weighted Regression - Hedonic Pricing Model for Private Highrise Property.
author:
  - name: Xu Pengtai
    url: https://www.linkedin.com/in/xupengtai/
date: 10-18-2021
output:
  distill::distill_article:
    self_contained: false
---


``` {r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      fig.retina = 3)
```

## Getting Started
create a list of R packages required

```{r echo=TRUE, eval=TRUE}
packages <- c('olsrr', 'GWmodel', 'tmap', 'sf', 'corrplot', 'tidyverse', 'ggpubr', 'spdep')
for (p in packages) {
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
```

## Import geospatial data into r environment
```{r echo=TRUE, eval=TRUE}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

- update CRS information with the correct ESPG code
```{r echo=TRUE, eval=TRUE}
mpsz_svy21 <- st_transform(mpsz, 3414)
```


## Import aspatial data into r environment
```{r echo=TRUE, eval=TRUE}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

```{r echo=TRUE, eval=TRUE}
glimpse(condo_resale)
```

- converting aspatial data frame into a sf object
```{r echo=TRUE, eval=TRUE}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
head(condo_resale.sf)
```

## EDA
```{r echo=TRUE, eval=TRUE}
AREA_SQM <- ggplot(data=condo_resale.sf, aes(x= `AREA_SQM`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

AGE <- ggplot(data=condo_resale.sf, aes(x= `AGE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CBD <- ggplot(data=condo_resale.sf, aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_CHILDCARE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_ELDERLYCARE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, aes(x= `PROX_URA_GROWTH_AREA`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, aes(x= `PROX_HAWKER_MARKET`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, aes(x= `PROX_KINDERGARTEN`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_MRT <- ggplot(data=condo_resale.sf, aes(x= `PROX_MRT`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PARK <- ggplot(data=condo_resale.sf, aes(x= `PROX_PARK`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, aes(x= `PROX_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, aes(x= `PROX_TOP_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, PROX_ELDERLYCARE, PROX_URA_GROWTH_AREA, PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT, PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  ncol = 3, nrow = 4)
```

## Drawing Statistical Point Map
```{r echo=TRUE, eval=TRUE}
tmap_mode("view")
tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(col = "SELLING_PRICE",
          alpha = 0.6,
          style="quantile") +
  tmap_options(check.and.fix = TRUE) +
  tm_view(set.zoom.limits = c(11,14))
tmap_mode("plot")
```


## simple linear regression
```{r echo=TRUE, eval=TRUE}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, data = condo_resale.sf)
summary(condo.slr)
```

- refer to scatter plot to check linearity
```{r echo=TRUE, eval=TRUE}
ggplot(data=condo_resale.sf,  
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) +
  geom_point() +
  geom_smooth(method = lm)
```

## multiple linear regression

- EDA
```{r echo=TRUE, eval=TRUE}
corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE",
         tl.pos = "td", tl.cex = 0.5, method = "number", type = "upper")
```

- regression
```{r echo=TRUE, eval=TRUE}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE  + PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET  + PROX_KINDERGARTEN  + PROX_MRT  + PROX_PARK  + PROX_PRIMARY_SCH + PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL  + PROX_SUPERMARKET + PROX_BUS_STOP  + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, data=condo_resale.sf)
summary(condo.mlr)
```

checks on linear regression assumptions

```{r echo=TRUE, eval=TRUE}
condo.mlr1 <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE  + PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK  + PROX_PRIMARY_SCH + PROX_SHOPPING_MALL  + PROX_BUS_STOP  + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, data=condo_resale.sf)
ols_regress(condo.mlr1)
```

- multicolinearity
```{r echo=TRUE, eval=TRUE}
ols_vif_tol(condo.mlr1)
```

- normality
```{r echo=TRUE, eval=TRUE}
ols_plot_resid_hist(condo.mlr1)
```

```{r echo=TRUE, eval=TRUE}
ols_test_normality(condo.mlr1)
```

- autocorrelation
```{r echo=TRUE, eval=TRUE}
condo_resale.res.sf <- cbind(condo_resale.sf, 
                        condo.mlr1$residuals) %>%
rename(`MLR_RES` = `condo.mlr1.residuals`)
condo_resale.sp <- as_Spatial(condo_resale.res.sf)
nb <- dnearneigh(coordinates(condo_resale.sp), 0, 1500, longlat = FALSE)
nb_lw <- nb2listw(nb, style = 'W')
lm.morantest(condo.mlr1, nb_lw)
```

## Building Hedonic Pricing Models using GWmodel

- fixed bandwith
```{r echo=TRUE, eval=TRUE}
bw.fixed <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE  + PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK  + PROX_PRIMARY_SCH + PROX_SHOPPING_MALL  + PROX_BUS_STOP  + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, data=condo_resale.sp, approach="CV", kernel="gaussian", adaptive=FALSE, longlat=FALSE)

gwr.fixed <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE  + PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK  + PROX_PRIMARY_SCH + PROX_SHOPPING_MALL  + PROX_BUS_STOP  + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, data=condo_resale.sp, bw=bw.fixed, kernel = 'gaussian', longlat = FALSE)


gwr.fixed
```

- adaptive
```{r echo=TRUE, eval=TRUE}
bw.adaptive <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE  + PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK  + PROX_PRIMARY_SCH + PROX_SHOPPING_MALL  + PROX_BUS_STOP  + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, data=condo_resale.sp, approach="CV", kernel="gaussian", adaptive=TRUE, longlat=FALSE)

gwr.adaptive <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE  + PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK  + PROX_PRIMARY_SCH + PROX_SHOPPING_MALL  + PROX_BUS_STOP  + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, data=condo_resale.sp, bw=bw.adaptive, kernel = 'gaussian', adaptive=TRUE, longlat = FALSE)

gwr.adaptive
```

## Visualisig local R2
```{r echo=TRUE, eval=TRUE}
condo_resale.sf.adaptive <- st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(crs=3414)
condo_resale.sf.adaptive.svy21 <- st_transform(condo_resale.sf.adaptive, 3414)
gwr.adaptive.output <- as.data.frame(gwr.adaptive$SDF)
condo_resale.sf.adaptive <- cbind(condo_resale.res.sf, as.matrix(gwr.adaptive.output))

tmap_mode("view")
tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))
```

- By URA planning region
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_svy21[mpsz_svy21$REGION_N=="CENTRAL REGION", ])+
  tm_polygons()+
tm_shape(condo_resale.sf.adaptive) + 
  tm_bubbles(col = "Local_R2",
           size = 0.15,
           border.col = "gray60",
           border.lwd = 1)
```


Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


