---
title: "hands-on-exercise-10"
description: |
  Spatial Interaction Models using GLM.
author:
  - name: Xu Pengtai
    url: https://www.linkedin.com/in/xupengtai/
date: 10-25-2021
output:
  distill::distill_article:
    self_contained: false
---


``` {r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      fig.retina = 3)
```

## Getting Started
create a list of R packages required

```{r echo=TRUE, eval=TRUE}
packages <- c('tmap', 'tidyverse',
             'sp', 'caret',
             'geojsonio', 'reshape2', 
             'broom')
for (p in packages) {
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}



library(sf)
library(stplanr)

```

## Import geospatial data into r environment
```{r echo=TRUE, eval=TRUE}
Aus <- geojson_read("https://www.dropbox.com/s/0fg80nzcxcsybii/GCCSA_2016_AUST_New.geojson?raw=1", what = "sp")
Ausdata <- Aus@data
AusSF <- st_as_sf(Aus) %>% 
  st_set_crs(4283) 
st_make_valid(AusSF)
```


Display data table
```{r echo=TRUE, eval=TRUE}
AusSF1 <- AusSF[order(AusSF$GCCSA_CODE),]
head(AusSF1, 10)
Aus <- as(AusSF1, "Spatial")
```

## Calculate a distance matrix
```{r echo=TRUE, eval=TRUE}
# re-project to projected coordinate system
AusProj <- spTransform(Aus,"+init=epsg:3112")

# compute distance matrix
dist <- spDists(AusProj)

# convert distance matrix into distance pair list
distPair <- melt(dist)

# convert unit of measurement from metres to km
distPair$value <- distPair$value / 1000
head(distPair, 10)
```

## Import interaction data into r environment
```{r echo=TRUE, eval=TRUE}
mdata <- read_csv("https://www.dropbox.com/s/wi3zxlq5pff1yda/AusMig2011.csv?raw=1",col_names = TRUE)
glimpse(mdata)
```

## Combine the imported migration data
```{r echo=TRUE, eval=TRUE}
mdata$FlowNoIntra <- ifelse(mdata$Orig_code == mdata$Dest_code,0,mdata$Flow)
mdata$offset <- ifelse(mdata$Orig_code == mdata$Dest_code,0.0000000001,1)
mdata$dist <- distPair$value 
mdata$dist <- ifelse(mdata$dist == 0,5,mdata$dist)
glimpse(mdata)
```

## Visualise with desire line
```{r echo=TRUE, eval=TRUE}
mdatasub <- mdata[mdata$Orig_code!=mdata$Dest_code,]
mdatasub_skinny <- mdatasub[,c(2,4,5)]
travel_network <- od2line(flow = mdatasub_skinny, 
                          zones = Aus)
travel_networkwgs <- spTransform(travel_network,"+init=epsg:4326" )
AusWGS <- spTransform(Aus,"+init=epsg:4326" )
w <- mdatasub_skinny$Flow / max(mdatasub_skinny$Flow) * 10
#plot(travel_networkwgs, lwd = w)
#plot(AusWGS, add=T)
```


## Build spatial ineraction models

### Unconstrained spatial interaction model
```{r echo=TRUE, eval=TRUE}
uncosim <- glm(Flow ~ log(vi1_origpop)+log(wj3_destmedinc)+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)

# fit the model
mdatasub$fitted <- fitted(uncosim)
mdatasubmat2 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "fitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "Flow", margins=c("Orig_code", "Dest_code"))
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `fitted`))+
  geom_point(color="black", fill="light blue")
postResample(mdatasub$Flow,mdatasub$fitted)
```

R squre value at 0.32. It seems that the uncontrained model failed to fit the empirical data well.

### Origin spatial interaction model
```{r echo=TRUE, eval=TRUE}
origSim <- glm(Flow ~ Orig_code+log(wj3_destmedinc)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
mdatasub$origSimFitted <- round(fitted(origSim),0)
mdatasubmat3 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "origSimFitted", margins=c("Orig_code", "Dest_code"))

# fit the model
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `origSimFitted`))+
  geom_point(color="black", fill="light blue")
postResample(mdatasub$Flow,mdatasub$origSimFitted)
```

R value improved substantially from 0.32 to 0.43 using the origin constrained model.

### Destination spatial interaction model
```{r echo=TRUE, eval=TRUE}
destSim <- glm(Flow ~ Dest_code+log(vi1_origpop)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
mdatasub$destSimFitted <- round(fitted(destSim),0)
mdatasubmat6 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "destSimFitted", margins=c("Orig_code", "Dest_code"))

# fit the model
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `destSimFitted`))+
  geom_point(color="black", fill="light blue")
postResample(mdatasub$Flow,mdatasub$destSimFitted)
```

R square improved to 0.65 using the destination constrained model.

### Doubly spatial interaction model
```{r echo=TRUE, eval=TRUE}
doubSim <- glm(Flow ~ Orig_code+Dest_code+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
mdatasub$doubsimFitted <- round(fitted(doubSim),0)
mdatasubmat7 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "doubsimFitted", margins=c("Orig_code", "Dest_code"))

# fit the model
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `doubsimFitted`))+
  geom_point(color="black", fill="light blue")
postResample(mdatasub$Flow,mdatasub$doubsimFitted)
```

R square now reaches 0.87. Doubly constrained model works the best in this case.


Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


