---
title: "take-home-exercise-2"
description: |
  Analysing airbnb listings in Singapore and the impact of covid-19 on them.
author:
  - name: Xu Pengtai
    url: https://www.linkedin.com/in/xupengtai/
date: 09-24-2021
output:
  distill::distill_article:
    self_contained: false
---


``` {r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      fig.retina = 3)
```

## Getting Started
- create a list of R packages required

```{r echo=TRUE, eval=TRUE}
packages <- c('maptools', 'sf', 'raster', 'spatstat', 'tidyverse', 'tmap', 'plotly', 'ggthemes')
for (p in packages) {
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
```

## Datasets

1. Airbnb: June 2019 and June 2021
2. onemapsgapi: hotels, tourist attractions, mrt services
3. sg outline

```{r echo=TRUE, eval=TRUE}
# sg outline
sg_sf <- st_read(dsn = 'data/subzone_sg',
                   layer = "MP14_SUBZONE_WEB_PL")
sg <- as_Spatial(sg_sf)
sg_sp <- as(sg, "SpatialPolygons")
sg_owin <- as(sg_sp, "owin")
plot(sg_owin)
```


## Section A: Airbnb Distribution in 2019
In this section, you are tasked to investigate if the distribution of Airbnb listings are affected by location factors such as near to existing hotels, MRT services and tourist attractions. For the purpose of this analysis, Airbnb listings as at June 2019 should be used.

1. Exploratory Spatial Data Analysis

Derive kernel density maps of Airbnb listings, hotels, MRT services, and tourist attractions.

```{r eval=TRUE}
# listing
setwd('/Users/pengtaixu/IS415_blog/')
abb19 <- read_csv("./_posts/2021-09-24-take-home-exercise-2/data/airbnb/listings_2019.csv")
abb19_sf <- st_as_sf(abb19, 
                    coords = c("longitude",
                               "latitude"),
                    crs = 4326) %>%
  st_transform(crs=3414) 
abb19 <- as_Spatial(abb19_sf)
abb19_sp <- as(abb19, "SpatialPoints")
abb19_ppp <- as(abb19_sp, "ppp")

abb19_ppp = abb19_ppp[sg_owin]

abb19_ppp.km <- rescale(abb19_ppp, 1000, "km")
kde_abb19.bw <- density(abb19_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_abb19.bw)
```


```{r}
# hotels
hotels <- read_csv("_posts/2021-09-24-take-home-exercise-2/data/OneMap_Data/hotels.csv")
hotels_sf <- st_as_sf(hotels, 
                    coords = c("Lng",
                               "Lat"),
                    crs = 4326) %>%
  st_transform(crs=3414) 
hotels <- as_Spatial(hotels_sf)
hotels_sp <- as(hotels, "SpatialPoints")
hotels_ppp <- as(hotels_sp, "ppp")

hotels_ppp = hotels_ppp[sg_owin]

hotels_ppp.km <- rescale(hotels_ppp, 1000, "km")
kde_hotels.bw <- density(hotels_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_hotels.bw)

# mrt
mrt_sf <- st_read(dsn = 'data/OneMap_Data/mrt',
                   layer = "MRTLRTStnPtt")
mrt <- read_csv("_posts/2021-09-24-take-home-exercise-2/data/OneMap_Data/mrt.csv")
mrt_sf <- st_as_sf(mrt, 
                    coords = c("Lng",
                               "Lat"),
                    crs = 4326) %>%
  st_transform(crs=3414) 
mrt <- as_Spatial(mrt_sf)
mrt_sp <- as(mrt, "SpatialPoints")
mrt_ppp <- as(mrt_sp, "ppp")

mrt_ppp = mrt_ppp[sg_owin]

mrt_ppp.km <- rescale(mrt_ppp, 1000, "km")
kde_mrt.bw <- density(mrt_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_mrt.bw)

# tourist attractions
tour <- read_csv("_posts/2021-09-24-take-home-exercise-2/data/OneMap_Data/tourism.csv")
tour <- tour[complete.cases(tour[,c("LONGTITUDE","LATITUDE")]), ] # remove na
tour_sf <- st_as_sf(tour, 
                    coords = c("LONGTITUDE",
                               "LATITUDE"),
                    crs = 4326) %>%
  st_transform(crs=3414) 
tour <- as_Spatial(tour_sf)
tour_sp <- as(tour, "SpatialPoints")
tour_ppp <- as(tour_sp, "ppp")

tour_ppp = tour_ppp[sg_owin]

tour_ppp.km <- rescale(tour_ppp, 1000, "km")
kde_tour.bw <- density(tour_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_tour.bw)

```

Using appropriate tmap functions, display the kernel density maps on openstreetmap of Singapore. Describe the spatial patterns revealed by the kernel density maps. Highlight the advantage of kernel density map over point map.

```{r eval=TRUE}
# listing
gridded_kde_abb19_bw <- as.SpatialGridDataFrame.im(kde_abb19.bw)
kde_abb19_bw_raster <- raster(gridded_kde_abb19_bw)
tm_shape(kde_abb19_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```


```{r}

# hotels
gridded_kde_hotels_bw <- as.SpatialGridDataFrame.im(kde_hotels.bw)
kde_hotels_bw_raster <- raster(gridded_kde_hotels_bw)
tm_shape(kde_hotels_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

# mrt
gridded_kde_mrt_bw <- as.SpatialGridDataFrame.im(kde_mrt.bw)
kde_mrt_bw_raster <- raster(gridded_kde_mrt_bw)
tm_shape(kde_mrt_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

# tourist attractions
gridded_kde_tour_bw <- as.SpatialGridDataFrame.im(kde_tour.bw)
kde_tour_bw_raster <- raster(gridded_kde_tour_bw)
tm_shape(kde_tour_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

```

Kernel density map gives additional information on the clustering of the points compared to simple point analysis. 


2. Second-order Spatial Point Patterns Analysis

With reference to the spatial point patterns observed in (1):

Formulate the null hypothesis and alternative hypothesis and select the confidence level.

Ho = The distribution of airbnb listings in June 2019 in Singapore are randomly distributed.

H1= The distribution of airbnb listings in June 2019 in Singapore are not randomly distributed.

The 95% confident interval will be used.


Perform the test by using appropriate Second order spatial point patterns analysis technique.
```{r eval = TRUE }
clarkevans.test(abb19_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

With reference to the analysis results, draw statistical conclusions.

With the p value = 0.01 , we reject the null hypothesis and conclude that the distribution of airbnb listings in June 2019 in Singapore are not randomly distributed. Factors involved could be distance to hotels/mrts/tourist attractions. In addition, we do see the overlap of dense areas in the kernel density map of airbnb listings and hotels/tourist attractions. 

## Section B: Impact of COVID-19
In this section, you are required to analyse the impact of COVID-19 on Airbnb business in Singapore by comparing Airbnb listings data on June 2019 and June 2021.

3. Exploratory Spatial Data Analysis

Derive kernel density maps of all Airbnb listings and Airbnb by room type as at June 2019 and June 2021,
```{r eval = TRUE}
## June 2019
unique(abb19$room_type)

# Private room
private19 <- abb19[abb19$room_type=="Private room",]
private19_sp <- as(private19, "SpatialPoints")
private19_ppp <- as(private19_sp, "ppp")

private19_ppp = private19_ppp[sg_owin]

private19_ppp.km <- rescale(private19_ppp, 1000, "km")
kde_private19.bw <- density(private19_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_private19.bw)

setwd('/Users/pengtaixu/IS415_blog/')
abb21 <- read_csv("_posts/2021-09-24-take-home-exercise-2/data/airbnb/listings_2021.csv")
abb21_sf <- st_as_sf(abb21, 
                    coords = c("longitude",
                               "latitude"),
                    crs = 4326) %>%
  st_transform(crs=3414) 
abb21 <- as_Spatial(abb21_sf)

unique(abb21$room_type)

# Private room
private21 <- abb21[abb21$room_type=="Private room",]
private21_sp <- as(private21, "SpatialPoints")
private21_ppp <- as(private21_sp, "ppp")

private21_ppp = private21_ppp[sg_owin]

private21_ppp.km <- rescale(private21_ppp, 1000, "km")
kde_private21.bw <- density(private21_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
#plot(kde_private21.bw)
```


```{r}
  
# Entire home/apt
entire19 <- abb19[abb19$room_type=="Entire home/apt",]
entire19_sp <- as(entire19, "SpatialPoints")
entire19_ppp <- as(entire19_sp, "ppp")

entire19_ppp = entire19_ppp[sg_owin]

entire19_ppp.km <- rescale(entire19_ppp, 1000, "km")
kde_entire19.bw <- density(entire19_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_entire19.bw)

# Shared room
shared19 <- abb19[abb19$room_type=="Shared room",]
shared19_sp <- as(shared19, "SpatialPoints")
shared19_ppp <- as(shared19_sp, "ppp")

shared19_ppp = shared19_ppp[sg_owin]

shared19_ppp.km <- rescale(shared19_ppp, 1000, "km")
kde_shared19.bw <- density(shared19_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_shared19.bw)

## June 2021
  
# Entire home/apt
entire21 <- abb21[abb21$room_type=="Entire home/apt",]
entire21_sp <- as(entire21, "SpatialPoints")
entire21_ppp <- as(entire21_sp, "ppp")

entire21_ppp = entire21_ppp[sg_owin]

entire21_ppp.km <- rescale(entire21_ppp, 1000, "km")
kde_entire21.bw <- density(entire21_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_entire21.bw)

# Shared room
shared21 <- abb21[abb21$room_type=="Shared room",]
shared21_sp <- as(shared21, "SpatialPoints")
shared21_ppp <- as(shared21_sp, "ppp")

shared21_ppp = shared21_ppp[sg_owin]

shared21_ppp.km <- rescale(shared21_ppp, 1000, "km")
kde_shared21.bw <- density(shared21_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_shared21.bw)

# Hotel room
hotel21 <- abb21[abb21$room_type=="Hotel room",]
hotel21_sp <- as(hotel21, "SpatialPoints")
hotel21_ppp <- as(hotel21_sp, "ppp")

hotel21_ppp = hotel21_ppp[sg_owin]

hotel21_ppp.km <- rescale(hotel21_ppp, 1000, "km")
kde_hotel21.bw <- density(hotel21_ppp.km,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
plot(kde_hotel21.bw)

```



Using appropriate tmap functions, display the kernel density maps on openstreetmap of Singapore,

```{r eval = TRUE}
## June 2019 

# Private room
kde_private19.bw
gridded_kde_private19_bw <- as.SpatialGridDataFrame.im(kde_private19.bw)
kde_private19_bw_raster <- raster(gridded_kde_private19_bw)
tm_shape(kde_private19_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

## June 2021

# Private room
kde_private21.bw
gridded_kde_private21_bw <- as.SpatialGridDataFrame.im(kde_private21.bw)
kde_private21_bw_raster <- raster(gridded_kde_private21_bw)
tm_shape(kde_private21_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

```


```{r}

# Entire home/apt
kde_entire19.bw
gridded_kde_entire19_bw <- as.SpatialGridDataFrame.im(kde_entire19.bw)
kde_entire19_bw_raster <- raster(gridded_kde_entire19_bw)
tm_shape(kde_entire19_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

# Shared room
kde_shared19.bw
gridded_kde_shared19_bw <- as.SpatialGridDataFrame.im(kde_shared19.bw)
kde_shared19_bw_raster <- raster(gridded_kde_shared19_bw)
tm_shape(kde_shared19_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)


## June 2021 

# Entire home/apt
kde_entire21.bw
gridded_kde_entire21_bw <- as.SpatialGridDataFrame.im(kde_entire21.bw)
kde_entire21_bw_raster <- raster(gridded_kde_entire21_bw)
tm_shape(kde_entire21_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

# Shared room
kde_shared21.bw
gridded_kde_shared21_bw <- as.SpatialGridDataFrame.im(kde_shared21.bw)
kde_shared21_bw_raster <- raster(gridded_kde_shared21_bw)
tm_shape(kde_shared21_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

# Hotel room
kde_hotel21.bw
gridded_kde_hotel21_bw <- as.SpatialGridDataFrame.im(kde_hotel21.bw)
kde_hotel21_bw_raster <- raster(gridded_kde_hotel21_bw)
tm_shape(kde_hotel21_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)


```



Describe the spatial patterns reveal by the kernel density maps.



4. Second-order Spatial Point Patterns Analysis

With reference to the spatial point patterns observed in (3):

Formulate the null hypothesis and alternative hypothesis and select the confidence level.
We can form null hypothesis on both 2019 and 2021 airbnb listings by room type. For example, we can form the following two set of hypothesis to have an understanding on the effect of covid-19 on the distribution of the listing of private rooms available on airbnb. 

Set 1:
Ho = The distribution of airbnb private room listings in June 2019 in Singapore are randomly distributed.
H1= The distribution of airbnb private listings in June 2019 in Singapore are not randomly distributed.
The 95% confident interval will be used.

```{r eval=TRUE}
clarkevans.test(private19_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

Set 2:
Ho = The distribution of airbnb private room listings in June 2021 in Singapore are randomly distributed.
H1= The distribution of airbnb private listings in June 2021 in Singapore are not randomly distributed.
The 95% confident interval will be used.

```{r eval=TRUE}
clarkevans.test(private21_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

With reference to the analysis results, draw statistical conclusions.
Set 1: R = 0.38104, p-value = 0.01
Set 2: R = 0.32079, p-value = 0.01
While both sets have the same p-value of 0.01, set 2 has a smaller Nearest Neighbour Ratio. This smaller value means that the private room listings on airbnb in 2021 
have got more clustered compared to that before covid-19 in 2019. 
We can compare the other room types in a similar manner. 



Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


