---
title: "take-home-exercise-1"
description: |
 Geospatial analytics on the development of covid-19 in Jarkarta, Indonesia.
author:
  - name: Xu Pengtai
    url: https://www.linkedin.com/in/xupengtai/
date: 09-04-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = TRUE,
  comment = "##",
  R.options = list(width = 60)
)
```

## Getting Started
- create a list of R packages required

```{r echo=TRUE, eval=TRUE}
packages <- c('sf', 'tidyverse', 'tmap')
for (p in packages) {
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
```


## Data Import
- aspatial: covid data by month
```{r echo=TRUE, eval=TRUE}
setwd("~/IS415_blog/_posts/2021-09-04-take-home-exercise-1")
library(readxl)
covid_monthly <- read_excel("data/aspatial/covid_monthly.xlsx")
```

- geospatial: DKI Jarkarta
```{r echo=TRUE, eval=TRUE, R.options=list(width = 70)}
jarkarta <- st_read("./data/geospatial/BATAS DESA DESEMBER 2019 DUKCAPIL DKI JAKARTA/BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA.shp")
jarkarta_sel = jarkarta[,1:9]

```

## Data Wrangling
- remove outer islands
```{r echo=TRUE, eval=TRUE}
jarkarta_main <- jarkarta_sel[!grepl("PULAU", jarkarta_sel$DESA),]
plot(st_geometry(jarkarta_main))
```

- left join: covid_monthly + jarkarta_main
```{r echo=TRUE, eval=TRUE}
names(covid_monthly)[names(covid_monthly) == 'ID_KEL'] <- 'KODE_DESA'
jarkarta_covid <- merge(x = covid_monthly, y = jarkarta_main, by = "KODE_DESA", all.x = TRUE)
jarkarta_covid <- st_as_sf(jarkarta_covid)
jarkarta_covid <- jarkarta_covid[!is.na(jarkarta_covid$OBJECT_ID),]
```

## Calculations
- calculate the cumulative confirmed cases per 10000 population
```{r echo=TRUE, eval=TRUE}
jarkarta_covid <- jarkarta_covid %>% 
   mutate(CASE_RATE =  POSITIF / (JUMLAH_PEN / 10000))

cumulative_death_by_month <- aggregate(jarkarta_covid$CASE_RATE, by=list(Category=jarkarta_covid$Date), FUN=mean)
names(cumulative_death_by_month) <- c("Date", "Cumulative Case Rates")
print(cumulative_death_by_month)
```

- calculate the cumulative death by month
```{r echo=TRUE, eval=TRUE}
jarkarta_covid <- jarkarta_covid %>% 
   mutate(DEATH_RATE = Meninggal / (JUMLAH_PEN / 10000))

cumulative_death_by_month <- aggregate(jarkarta_covid$DEATH_RATE, by=list(Category=jarkarta_covid$Date), FUN=mean)
names(cumulative_death_by_month) <- c("Date", "Cumulative Death Rates")
print(cumulative_death_by_month)
```

## Plots

- thematic mapping: cumulative confirmed cases rate
```{r echo=TRUE, eval=TRUE}

tm_shape(jarkarta_covid) +
    tm_polygons("CASE_RATE") +
    tm_facets(by = "Date")
```

- thematic mapping: cumulative death rate
```{r echo=TRUE, eval=TRUE}

tm_shape(jarkarta_covid) +
    tm_polygons("DEATH_RATE") +
    tm_facets(by = "Date")
```

- analytical mapping: cumulative confirmed cases rate
```{r echo=TRUE, eval=TRUE}

get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
# percent map 
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0, .01, .1, .5, .9, .99, 1)
  var <- get.var(vnam,df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels = c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right", "bottom")) +
  tm_facets(by = "Date")
}

percentmap("CASE_RATE", jarkarta_covid)
```


- analytical mapping: cumulative death rate
```{r echo=TRUE, eval=TRUE}

percentmap("DEATH_RATE", jarkarta_covid)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


