---
title: "hands-on-exercise-7"
description: |
  Measures of Spatial Autocorrelation.
author:
  - name: Xu Pengtai
    url: https://www.linkedin.com/in/xupengtai/
date: 09-27-2021
output:
  distill::distill_article:
    self_contained: false
---


``` {r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      fig.retina = 3)
```

## Getting Started
create a list of R packages required

```{r echo=TRUE, eval=TRUE}
packages <- c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages) {
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
```

## Import shapefile into r environment
```{r echo=TRUE, eval=TRUE}
hunan <- st_read(dsn = "data/shapefile", 
                 layer = "Hunan")
```

## Import csv file into r environment
```{r echo=TRUE, eval=TRUE}
hunan2012 <- read_csv("data/attribute/Hunan_2012.csv")
```

## Performing relational join
```{r echo=TRUE, eval=TRUE}
hunan <- left_join(hunan,hunan2012)
```

## Visualising Regional Development Indicator
```{r echo=TRUE, eval=TRUE}
equal <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")

quantile <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

## Computing Contiguity Spatial Weights
```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

### Row-standardised weights matrix
```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

## Global Spatial Autocorrelation: Moran's I
```{r}
moran.test(hunan$GDPPC, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)

```

### Computing Monte Carlo Moran's I
```{r}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC, 
                listw=rswm_q, 
                nsim=999, 
                zero.policy = TRUE, 
                na.action=na.omit)
bperm
```

## Visualisinng Monte Carlo Moran's I 
```{r}
summary(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
hist(bperm$res, 
     freq=TRUE, 
     breaks=20, 
     xlab="Simulated Moran's I")
abline(v=0, 
       col="red") 
```

## Global Correlogram 

show how correlated are pairs of spatial observations when you increase the distance (lag) between them

```{r}
MI_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="I", 
                          style="W")
plot(MI_corr)
```

```{r}
print(MI_corr)
```

## Cluster and Outlier Analysis

### Computing Local Moran's I
```{r}
fips <- order(hunan$County)
localMI <- localmoran(hunan$GDPPC, rswm_q)
head(localMI)
```

### Mapping Local Moran's I
```{r}
hunan.localMI <- cbind(hunan,localMI) %>%
  rename(Pr.Ii = Pr.z....E.Ii..)
```

Mapping local Moran’s I values
```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty",
          palette = "RdBu",
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)
```

Mapping local Moran’s I p-values
```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)
```

## Creating a LISA Cluster Map

- Moran Scatterplot: an illustration of the relationship between the values of the chosen attribute at each location and the average value of the same attribute at neighboring locations.
```{r}
nci <- moran.plot(hunan$GDPPC, rswm_q,
                  labels=as.character(hunan$County), 
                  xlab="GDPPC 2012", 
                  ylab="Spatially Lag GDPPC 2012")
```

Plotting Moran scatterplot with standardised variable

Centering is done by subtracting the mean (omitting NAs) the corresponding columns, and scaling is done by dividing the (centered) variable by their standard deviations.
```{r}
hunan$Z.GDPPC <- scale(hunan$GDPPC) %>% as.vector 
nci2 <- moran.plot(hunan$Z.GDPPC, rswm_q,
                   labels=as.character(hunan$County),
                   xlab="z-GDPPC 2012", 
                   ylab="Spatially Lag z-GDPPC 2012")
```

Preparing LISA map classes
```{r}
quadrant <- vector(mode="numeric",length=nrow(localMI))
DV <- hunan$GDPPC - mean(hunan$GDPPC)     
C_mI <- localMI[,1] - mean(localMI[,1])    
signif <- 0.05       
quadrant[DV >0 & C_mI>0] <- 4      
quadrant[DV <0 & C_mI<0] <- 1      
quadrant[DV <0 & C_mI>0] <- 2
quadrant[DV >0 & C_mI<0] <- 3
quadrant[localMI[,5]>signif] <- 0
```

- plot both the local Moran’s I values map and its corresponding p-values map next to each other.
```{r}
gdppc <- qtm(hunan, "GDPPC")

hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

LISAmap <- tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)

tmap_arrange(gdppc, LISAmap, asp=1, ncol=2)
```

## Hot Spot and Cold Spot Area Analysis

### Deriving the centroid
```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
coords <- cbind(longitude, latitude)
```

### Determinig the cut-off distance
```{r}
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```
set cut-off distance as 62 to ensure that each county has at least 1 neighbour

### Computing fixed distance weight matrix
```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm62_lw <- nb2listw(wm_d62, style = 'B')
summary(wm62_lw)
```

### Computing adaptive distance weight matrix
```{r}
knn <- knn2nb(knearneigh(coords, k=8))
knn_lw <- nb2listw(knn, style = 'B')
summary(knn_lw)
```

### Computing Gi statistics and Mapping Gi values

Using fixed distance
```{r}
fips <- order(hunan$County)
gi.fixed <- localG(hunan$GDPPC, wm62_lw)
hunan.gi <- cbind(hunan, as.matrix(gi.fixed)) %>%
  rename(gstat_fixed = as.matrix.gi.fixed.)

```

```{r}
gdppc <- qtm(hunan, "GDPPC")

Gimap <-tm_shape(hunan.gi) +
  tm_fill(col = "gstat_fixed", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)
```

Using adaptive distance
```{r}
fips <- order(hunan$County)
gi.adaptive <- localG(hunan$GDPPC, knn_lw)
hunan.gi <- cbind(hunan, as.matrix(gi.adaptive)) %>%
  rename(gstat_adaptive = as.matrix.gi.adaptive.)
gdppc <- qtm(hunan, "GDPPC")

Gimap <- tm_shape(hunan.gi) +
  tm_fill(col = "gstat_adaptive", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)

```


Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


